---
title: "localhierarchy"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# FPET2

R package for fitting Bayesian transition models to produce estimates
and projections of family planning indicators. This work was supported,
in whole or in part, by the Bill & Melinda Gates Foundation (INV-00844).

# Installation

Dependencies

-   `cmdstanr`: Instructions for installing `cmdstanr` are available in
    their [Getting
    started](https://mc-stan.org/cmdstanr/articles/cmdstanr.html) guide.
-   R package `localhierarchy`, available at
    [github.com/AlkemaLab/localhierarchy](github.com/AlkemaLab/localhierarchy).
    You can install it
    using`remotes::install_github("AlkemaLab/localhierarchy")`.

Install `fpet2` from Github:

`remotes::install_github("AlkemaLab/fpet2")`

# Information

The modeling approach is described in Alkema et al. (2024). This
approach builds off collaborative research with the United Nations
Population Division (UNPD). References to earlier and related work are
included below.

The national-level data and population estimates by marital status are
compiled and published by the United Nations Population Division (UNPD).
Track20 compiled a database of subnational data.

References that describe the current approach:

-   FPET overview: L. Alkema, H. Susmann, E. Ray, S. Mooney, N.
    Cahill, K. Bietsch, A. Jayachandran, et al. (2024). Statistical
    Demography Meets Ministry of Health: The Case of the Family Planning
    Estimation Tool. See <https://doi.org/10.48550/arXiv.2501.00007>.

-   Use of service statistics data (Estimated modern use): S. Mooney, L.
    Alkema, E. Sonneveldt, K. Bietsch, J. Williamson, N. Cahill (2024).
    Enhancing the Use of Family Planning Service Statistics Using a
    Bayesian Modelling Approach to Inform Estimates of Modern
    Contraceptive Use in Low- and Middle-Income Countries. See
    <https://doi.org/10.48550/arXiv.2412.08606>.

References that describe earlier approaches:

-   Alkema, L. et al. (2013). National, regional and global rates and
    trends in contraceptive prevalence and unmet need for family
    planning between 1990 and 2015: a systematic and comprehensive
    analysis. The Lancet. Available at
    [http://www.thelancet.com/journals/lancet/article/PIIS0140-6736(12)62204-1/abstract](http://www.thelancet.com/journals/lancet/article/PIIS0140-6736(12)62204-1/abstract){.uri}.

-   Cahill et al., (2017). Modern contraceptive use, unmet need, and
    demand satisfied among women of reproductive age who are married or
    in a union in the focus countries of the Family Planning 2020
    initiative: a systematic analysis using the Family Planning
    Estimation Tool. The Lancet. Available at
    [http://www.thelancet.com/journals/lancet/article/PIIS0140-6736(17)33104-5/fulltext](http://www.thelancet.com/journals/lancet/article/PIIS0140-6736(17)33104-5/fulltext){.uri}.

-   Kantorova et al. (2020). Estimating progress towardsmeeting women’s
    contraceptive needs in 185 countries: A Bayesian hierarchical
    modellingstudy. PLOS Medicine, 17(2):e1003026. Available at
    <https://journals.plos.org/plosmedicine/article?id=10.1371/journal.pmed.1003026>.

-   Wheldon et al. (2018). Methods for estimating and projecting key
    family planning indicators among all women of reproductive age.
    United Nations, Department of Economic and Social Affairs,
    Population Division, Technical Paper No. 2. New York: United
    Nations. Available at
    <https://www.un.org/en/development/desa/population/publications/pdf/technical/TP2018-2.pdf>.

The model is written in R and Stan. For any questions, feedback or
suggestions, please contact Leontine at lalkema(at)umass.edu.

# Citation

Please cite as follows:

```{r}
citation("fpet2")
```
